\documentclass[addpoints]{exam}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{mdframed}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amssymb}

\marksnotpoints
\pointsinrightmargin
\bracketedpoints
\printanswers
\renewcommand\qedsymbol{$\blacksquare$}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=blue,
  pdfpagemode=FullScreen,
}

\urlstyle{same}

\pagestyle{headandfoot}
\firstpageheadrule
\runningheadrule
\firstpageheader{Pre Calc Prep}{Induction PSET}{Shah}
\runningheader{Induction Problem Set}{}{Shah}
\firstpagefooter{}{}{}
\runningfooter{ }{\thepage}{ }

\begin{document}
\vspace{1in}

\noindent\makebox[\textwidth]{Name:\enspace\hrulefill}

\vspace{.1in}

\begin{center}
	\fbox{\fbox{\parbox{5.5in}{\centering
				Answer the questions in the spaces provided on the following pages.  If you run out of room for an answer, continue on the back of the page. Show \textbf{all} your work!}}}
\end{center}

\begin{questions}
	\question Prove the following using proof by induction
	\begin{center}
		For all positive integers, $n$, $\displaystyle\,1^2 + 2^2 + 3^2 + \cdots + n^2 = \frac{n(n+1)(2n+1)}{6}$
	\end{center}
	\begin{solution}[\stretch{1}]
		\begin{proof}
			Let $P(n)$ be the formula \[\displaystyle\,1^2 + 2^2 + 3^2 + \cdots + n^2 = \frac{n(n+1)(2n+1)}{6}\]
			\textbf{Show that $P(1)$ is true:} To show that $P(1)$ is true we will show that LHS and the RHS are equivalent. The LHS is simply $1^2=1$ and the RHS is $\displaystyle\, \frac{1(1+1)(2(1)+1)}{6} = \frac{1(2)(3)}{6} = 1$ and so the LHS and RHS equal the same quantity and thus must be equivalent, proving the truthfullness of $P(1)$
			\newline\textbf{Show that if $P(k)$ is true, so is $P(k+1)$:} First, we will assume that $P(k)$ is true or that
			\[
				1^2 + 2^2 + 3^2 + \cdots + k^2 = \frac{k(k+1)(2k+1)}{6}
			\]
			Next, we will need to show that $P(k+1)$ is true or that
			\[
				1^2 + 2^2 + 3^2 + \cdots + (k+1)^2 = \frac{(k+1)\left[\left(k+1\right)+1\right]\left[2\left(k+1\right)+1\right]}{6} = \frac{(k+1)(k+2)(2k+3)}{6}
			\]
			We can rewrite the LHS in the following way
			\begin{align*}
				1^2 + 2^2 + 3^2 + \cdots + k^2 + (k+1)^2 & = \frac{k(k+1)(2k+1)}{6} + (k+1)^2                           \\
				                                         & = \frac{k(k+1)(2k+1)}{6} + \frac{6(k+1)^2}{6}                \\
				                                         & = \frac{k(k+1)(2k+1)}{6} + \frac{6(k^2 + 2k + 1)}{6}         \\
				                                         & = \frac{(k^2+k)(2k+1)}{6} + \frac{6k^2 + 12k + 6}{6}         \\
				                                         & = \frac{2k^3 + 2k^2 + k^2 + k}{6} + \frac{6k^2 + 12k + 6}{6} \\
				                                         & = \frac{2k^3 + 9k^2 + 13k + 6}{6}
			\end{align*}
			The RHS can be expanded to give
			\begin{align*}
				\frac{(k+1)(k+2)(2k+3)}{6} & = \frac{(k^2 + 3k + 2)(2k+3)}{6}             \\
				                           & = \frac{2k^3 + 3k^2 + 6k^2 + 9k + 4k + 6}{6} \\
				                           & = \frac{2k^3 + 9k^2 + 13k + 6}{6}
			\end{align*}
			Now our LHS and RHS equal the same quantity and must be equivalent, thus our statement holds true through proof by induction.
		\end{proof}
	\end{solution}

	\newpage

	\question Prove the following using proof by induction
	\begin{center}
		For all integers $n \geq 1$, $\displaystyle\, 1 + 3 + 5 + \cdots + (2n-1) = n^2$
	\end{center}
	\begin{solution}[\stretch{1}]
		\begin{proof}
			Let $P(n)$ be the formula
			\[
				1 + 3 + 5 + \cdots + (2n-1) = n^2
			\]
			\textbf{Show that $P(1)$ is true:} To show that $P(1)$ is true, we will need to show that the LHS and RHS are equivalent. Well, the LHS is $2(1)-1=1$ and the RHS is $1^2=1$ and so both sides equal the same quantity and thus are equivalent, proving the truthfullness of $P(1)$
			\newline\textbf{Show that if $P(k)$ is true, so is $P(k+1)$:} We will assume that $P(k)$ is true or that
			\[
				1 + 3 + 5 + \cdots + (2k-1) = k^2
			\]
			and we must show that $P(k+1)$ is true, or that
			\[
				1 + 3 + 5 + \cdots + \left(2\left(k+1\right)-1\right) = \left(k+1\right)^2
			\]
			We can begin with the LHS as follows,
			\begin{align*}
				1 + 3 + 5 + \cdots + (2k + 1) & = 1 + 3 + 5 + \cdots + (2k - 1) + (2k + 1) \\
				                              & = k^2 + (2k + 1)                           \\
				                              & = (k+1)(k+1)                               \\
				                              & = \left(k+1\right)^2
			\end{align*}
			And so, our LHS and RHS equal the same quantity and thus are equivalent, proving our statement through proof by induction.
		\end{proof}
	\end{solution}

	\question Prove the following using proof by induction
	\begin{center}
		For all integers $n \geq 1$, $\displaystyle\, 2 + 4 + 6 + \cdots + 2n = n^2 + n$
	\end{center}
	\begin{solution}[\stretch{1}]
		\begin{proof}
			Let $P(n)$ be the formula
			\[
				2 + 4 + 6 + \cdots + 2n = n^2 + n
			\]
			\textbf{Show that $P(1)$ is true:} To show that $P(1)$ is true we will show that the LHS and the RHS are equivalent. The LHS is $2(1)=2$ and the RHS is $1^2 + 1 = 2$ and so, since the LHS and RHS are equivalent, $P(1)$ holds true.
			\newline\textbf{Show that if $P(k)$ is true, so is $P(k+1)$:} We will assume that $P(k)$ is true or that
			\[
				2 + 4 + 6 + \cdots + 2k = k^2 + k
			\]
			and we will show that $P(k+1)$ is true or that
			\[
				2 + 4 + 6 + \cdots + 2(k+1) = (k+1)^2 + k = k^2 + 2k + 1 + k = k^2 + 3k + 1
			\]
			We can begin with the LHS to show
			\begin{align*}
				2 + 4 + 6 + \cdots 2(k+1) & = 2 + 4 + 6 + \cdots + 2k + 2(k+1) \\
				                          & = (k^2 + k) + 2(k+1)               \\
				                          & = k^2 + k + 2k + 1                 \\
				                          & = k^2 + 3k + 1
			\end{align*}
			Which is the RHS of $P(k+1)$ and so our LHS and RHS are equivalent, proving the validity of our statement through proof by induction.
		\end{proof}
	\end{solution}

	\newpage

	\question \textit{(Challenge)} Prove the following using proof by induction
	\begin{center}
		For all integers $n \geq 1$, $\displaystyle\,\sum\limits_{i=1}^{n} i\left(i\textbf{!}\right) = (n+1)\textbf{!} - 1$
	\end{center}
	\begin{solution}[\stretch{1}]
		\begin{proof}
			Let $P(n)$ be the formula
			\[
				\sum\limits_{i=1}^{n} i\left(i\textbf{!}\right) = \left(n+1\right)\textbf{!} - 1
			\]
			\textbf{Show that $P(1)$ is true:} To show that $P(1)$ is true we will show that the LHS and RHS are equivalent.
			\newline
			\begin{minipage}[t]{0.45\linewidth}
				\begin{center}
					\underline{LHS}
				\end{center}
				\[
					\sum\limits_{i=1}^{1} i\left(i\textbf{!}\right) = 1(1\textbf{!}) = 1(1) = 1
				\]
			\end{minipage}
			\hfill
			\begin{minipage}[t]{0.45\linewidth}
				\begin{center}
					\underline{RHS}
				\end{center}
				\[
					\left(1+1\right)\textbf{!} - 1 = 2\textbf{!} - 1 = 2 - 1 = 1
				\]
			\end{minipage}
			\newline And so, our LHS and RHS are equivalent, thus proving $P(1)$
			\newline\textbf{Show that if $P(k)$ is true, so is $P(k+1)$:} First, we will assume that $P(k)$ is true or that
			\[
				\sum\limits_{i=1}^{k} i\left(i\textbf{!}\right) = \left(k+1\right)\textbf{!} - 1
			\]
			and we will need to show that $P(k+1)$ is true or that
			\[
				\sum\limits_{i=1}^{k+1} i\left(i\textbf{!}\right) = \left[\left(k+1\right)+1\right]\textbf{!}-1 = \left(k+2\right)\textbf{!} - 1
			\]
			We can begin with the LHS as follows,
			\begin{align*}
				\sum\limits_{i=1}^{k+1} i\left(i\textbf{!}\right) & = \left(k+1\right)\left[\left(k+1\right)\textbf{!}\right] + \sum\limits_{i=1}^{k} i\left(i\textbf{!}\right)                              \\
				                                                  & = \left(k+1\right)\left(k+1\right)\textbf{!} + \left[\left(k+1\right)\textbf{!} - 1\right]                                               \\
				                                                  & = \left[\left(k+1\right)\left(k+1\right)\textbf{!} + \left(k+1\right)\textbf{!}\right] - 1 \tag{by the associative property of addition} \\
				                                                  & = \left(k+1\right)\textbf{!}\left[\left(k+1\right) + 1\right] - 1 \tag{by factoring}                                                     \\
				                                                  & = \left(k+1\right)\textbf{!}\left(k+2\right) - 1                                                                                         \\
				                                                  & = \left(k+2\right)\textbf{!} - 1 \tag{by properties of factorials}
			\end{align*}
			And so, our LHS is equivalent to the same quantity as the RHS and is thus equal to the RHS, proving our statement through induction.
		\end{proof}
	\end{solution}
	\newpage
	\question \textit{(\textbf{Mega} Challenge)} Prove the following using proof by induction
	\begin{center}
		For all integers $n \geq 1$, $3$ is a factor of $4^n - 1$
	\end{center}
	\begin{solution}[\stretch{1}]
		\begin{proof}
			Let $P(n)$ be the statement that $3$ is a factor of $4^n - 1$
			\newline\textbf{Show that $P(1)$ is true:} To show that $P(1)$ is true we will show that $3$ is a factor of $4^1 - 1$. $4^1 - 1 = 3$ and $3/3 = 1$ which is an integer, thus $3$ is a factor of $4^1-1$ and $P(1)$ is valid.
			\newline\textbf{Show that if $P(k)$ is true, so is $P(k+1)$:} We will assume that $P(k)$ is true or that
			\[
				\frac{4^k-1}{3} \in \mathbb{Z}
			\]
			and we will need to show that
			\[
				\frac{4^{k+1}-1}{3} \in \mathbb{Z}
			\]
			We can do this by manipulating the LHS
			\begin{align*}
				4^{k+1} - 1 & = 4^{k+1} - 4^k + 4^k - 1 \\
				            & = 4^k * 4 - 4^k + 4^k - 1 \\
				            & = 4^k(4-1) + (4^k - 1)    \\
				            & = 4^k(3) + (4^k - 1)
			\end{align*}
			Since we already know that $4^k-1$ has a factor of $3$ we can conclude that $4^{k+1}$ must also have a factor of $3$ since $\displaystyle\, \frac{4^k(3)}{3} = 4^k \in \mathbb{Z}$ and so adding another factor of $3$ will still give a factor of $3$. Thus, our statement holds true through proof by induction.
		\end{proof}
	\end{solution}
\end{questions}
\end{document}
